---
title: 'Homework #2'
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example question and answer
0\. What is the smallest positive odd integer? Storage it in a variable named `a0` and print it.

**Example answer**:
In the RMarkDown file, type in
<pre>```{r}
a0 <- 1
a0
```</pre>
The knitted html file will show:
```{r}
a0 <- 1
a0
```

## R commands
1. A student's exam grades in a class were 95, 65, 70, and 87. The course policy was to drop the lowest exam grade. Find the average exam grade after dropping the lowest. Store the answer in a variable named `ave` and print.
2. Create a vector of numbers (with length 21) to show whether the years of 2000, 2001, ..., 2020 are leap years. An entry of the vector should be `1` if the corresponding year is a leap year, and `0` otherwise. A year is a leap year here if it is divisible by 4. Store the answer in a variable named `leap` and print.
3. Calculate the BMI of Chris and Mary, who are siblings. Chris is 1.7 meters tall and 80 kilograms in weight; Mary is 1.65 meters tall and 49 kilograms in weight. The formula for BMI is `BMI = weight^2 (in kg) / height (in meter)`. Store the answers in a vector of length 2 named `bmi` and print.

## Cars and mileage
For this part we will use the `mpg` data set, you can access the data by first installing (using `install.packages('ggplot2')`) and then loading the package `ggplot2` (using `library(ggplot2)`) into your R session. The `mpg` data includes information about fuel economy of popular car models
for the years 1999 and 2008 as collected by the US Environmental Protection Agency. An answer should include a short paragraph of text for the explanation to a (human) reader, together with appropriate R code, outputs, and graphs.

1. What are the 5 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 5 rows) containing all information of these five cars in a variable named `best5` and print. (Hint: Here are a few things you are free to define. E.g., which variable to use to characterize "fuel economy", and what is a "model". Please state your definitions.)
2. How many different SUV models are included in this dataset? Storage the answer in a variable named `numSUV` and print.
3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible? Explain your findings.
4. Which manufacturer produces cars with higher fuel efficiency, Chevrolet or Honda? Use graphs *and* summary statistics to answer this question. Explain your findings.
