---
title: 'Homework #3'
author: "Your Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tips at Restaurants

In this homework we will use the `tips` data set. The dataset contains tip information and is collected by one waiter, who recorded each tip he received while working in one restaurant over the course of a season. This data set is part of the `reshape2` package. You can load the data set by executing the command:
```{r}
data(tips, package="reshape2")
```
If you do not have available the package `reshape2`, issue `install.packages('reshape2')` to install. See `?tips` for a description of all of the variables. 

Include at least one `ggplot` visualization in each of your answer, and describe what you find.
<!-- **Submission instructions**: Create a folder named `ds202_hw3`, and name the RMarkDown file `hw3.Rmd`  which should include your solutions under this folder. For submission, create a GitHub repository named `ds202_hw3` under your GitHub account, and push both `hw3.Rmd` and the knitted `hw3.html` before the deadline. I will assume you use the same GitHub username as for your HW2 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. This homework is **due on Feb 19** before class. -->

1. Explore and explain what is the relationship between the total bill and tip amount. In your visualization, add a straight line with intercept at 0 and slope equal to 0.15, which stands for the amount of tip according to a 15% tip rate. Look at `?geom_abline`, and in particular the examples.
2. Describe at least two types of anomalies in the previous plot. What do they mean?
3. Visualize the number of customers the waiter has on each day of the week. Make sure the days are ordered from Thursday to Sunday. What pattern do you find?
4. Introduce a variable `tiprate` into the data set which stands for the rate of tips. How does gender affect the distribution of tip rates?
5. How do tip rates vary on different days of the week?
6. How does smoking behavior and gender of the person who pays impact the relationship between tip and total bill? Find a visualization that incorporates all four variables. Interpret the result.
7. Use ggplot2 to find a graphical summary of the relationship between day of the week and gender of the person paying the bill. What can you say about this relationship?
