---
title: 'Homework #3'
author: "Your Name"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## House Price in Ames

In this homework we will use the `ames` dataset. The dataset contains house sales information in Ames between 2006 and 2010. The original dataset was compiled by Dean De Cock, and the version we are using is extracted from [Kaggle](https://www.kaggle.com/c/house-prices-advanced-regression-techniques/overview).

The dataset is included in our `classdata` library. If you have not installed this, issue in R
```{r,eval=FALSE}
devtools::install_github('xdaiISU/classdata')
```

To load the dataset, 
```{r,eval=FALSE}
library(classdata)
data(ames)
```
See `?ames` for the description on the dataset.

Include at least one `ggplot` visualization in each of your answer except for Question 2, and describe what you find.
<!-- **Submission instructions**: Create a folder named `ds202_hw3`, and name the RMarkDown file `hw3.Rmd`  which should include your solutions under this folder. For submission, create a GitHub repository named `ds202_hw3` under your GitHub account, and push both `hw3.Rmd` and the knitted `hw3.html` before the deadline. I will assume you use the same GitHub username as for your HW2 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. This homework is **due on Feb 19** before class. -->

1. Explore and explain what is the relationship between the sale price (y) and living area size (x). In your visualization, add a straight line with intercept at 20000 and slope equal to 100. Look at `?geom_abline`, and in particular the examples. Remark on any patterns you find and the straight line.
2. Describe at least two types of anomalies in the previous plot. What do they mean?
3. Visualize the number of sales by the overall condition. Make sure the conditions are ordered from the best to the worst. What do you find?
4. Introduce a variable `houseAge` into the data set which stands for the age of the house at the time when it was sold. How does the age of the house affect the sale price? 
5. Were there more sales of nice houses or poor houses? Make sure to state your criterion for a house to be "nice" or otherwise "poor". Choose an appropriate graph type.
6. How do the neighborhood and the slope of property affect the relationship between sale price and year built? Focus on the neighborhoods of Brookside, Clear Creek, College Creek, and Somerset only. Find a visualization that incorporates all four variables. Interpret the result.
7. Create a side-by-side histogram showing the sales price for the different types of buildings. Comment on the result.
