---
title: "DS 202 - lab #4: Joining and Tidying"
author: "Xiongtao Dai"
ratio: 16x10
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE, cache=TRUE)
```

## Outline

- Optionally, work in assigned pairs
- Create a repository (one per team) & give access to your teammate
- Explore the Cyclone football data. Use `dplyr` to manipulate data frames and `tidyr` to tidy data up. 
- Create an R markdown file that contains all code to read the data, create the plots, and your answers.
- Teammates will review and revise each other's work.
- Submit!


## Submission Requirements

- A team member should create a GitHub repository called `ds202_lab4` which includes the group work.
- Each team member should create at least one commit with message "Review and Revision".
- The team will collaborate on `team.Rmd` to answer all the questions, which will be graded. Also include your individual exploration in either `individual1.Rmd` or `individual2.Rmd`. **Even if you are working alone, you must include a team.Rmd file which contains your report to be graded**.
- The correct folder structure is as below. The folder `ds202_lab4` is a git repository.

    ```
    ds202_lab4
    |-- individual1.Rmd
    |-- individual2.Rmd
    |-- team.Rmd
    |-- *.html
    |-- Data files if needed...
    ```

## Background
An excerpt of the individual player data of the Cyclone football team is extracted from the ISU Athletics website and is available on the course website. Explore the [2019 data file](../slides/cyclonesFootball2019.xlsx).

Some abbreviations for the column names can be found in the Glossaries sheet in the Excel file.

**References:**

- [Iowa State University Athelete](https://cyclones.com/index.aspx)
- [A Beginner's Guide to American Football](https://www.youtube.com/watch?v=3t6hM5tRlfA) for the rules of American football.

## Loading Data

The Cyclone football data involves multiple Excel sheets and is in a somewhat messy form. The following code loads the `Defensive` sheet into R. For this lab, also import the `Offensive` and the `Biography` sheets.
```{r}
library(readxl)
defense <- read_excel('../slides/cyclonesFootball2019.xlsx', sheet='Defensive')
str(defense)
```

```{r,echo=FALSE,results='hide'}
offense <- read_excel('../slides/cyclonesFootball2019.xlsx', sheet='Offensive')
bio <- read_excel('../slides/cyclonesFootball2019.xlsx', sheet='Biography')
```

## Part one: Cleaning data
Although the three Excel sheets are loaded into R, they are not useful before some cleaning. Manipulate the three imported data frames, so that in the cleaned dataset: 

1. The player names and opponent names are factors.
2. All offensive and defensive statistics are numerical. The `Weight` column in biogrpahy is also numeric. (Optional: Learn to mutate multiple columns systematically using the `?across` function in `dplyr)
3. Change the `Height` column in biogrpahy into numeric. (Hint: Use a unit that works well)

By cleaning the three data frames in R, at the end of this step you should have the following three clean data frames: `defClean`, `offClean`, and `bioClean`. Print the structures of these data frames.

## Part two: Tidying
For the following questions, work on the cleaned data frames `defClean`, `offClean`, and `bioClean`.

1. Reformat the `defClean` data frame into a tidy format using `pivot_longer`. The type of statistic (Tackles_Solo, Tackles_ASST, etc) is added as a new key column named `stat`. 
2. Compare the distributions of the defensive statistics. What defensive skills are rare?
3. Did ISU have better defense against Iowa or Notre Dame? Answer this question by creating a scatterplot with x- and y-axes being the number of solo tackles (of each player). A large number of solo tackles is an indicator of good defense. 

  Now turn our attention to the player biography.

4. Separate the `Hometown` column into two columns, namely the city and the state. Print the first few rows of these two columns. (Hint: look at the `sep=` argument of `separate`)
5. How many players are there from each state?
```{r,echo=FALSE,results='hide',message=FALSE}
library(tidyverse)
a <- bio %>% separate(Hometown, c('City', 'State'), sep = ', ')
```

## Part three: Joining data frames

1. Find which states do the ISU defensive and offensive players come from. Create an appropriate summary.
<!-- 2. Is there a relationship between the player statistics and their positions? -->
3. How does the performance of quarterback Brock Purdy relate to the performance of the ISU team?
4. Which ISU player(s) made big progress over the last year? Compare the 2019 and the 2018 data [here](../slides/cyclonesFootball2018.xlsx). 


## Tips
- You may want to find one or more suitable statistic to quantify the performance of a player.
- Agree on a method to clean the dataset with your teammate. Both teammates can then work on the same cleaned version for downstream analysis.
- Plan ahead to prevent merge conflict: Teammates should work on different sections of `team.Rmd`. If they need to modify the same section, do it in turn rather than concurrently. 