---
title: "DS 202 - lab #5: Drunk Driving and Fatal Accidents "
ratio: 16x10
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE, cache=TRUE)
```

## Outline

- Optionally, form a team with another student
- Create a repository (one per team) & give access to your teammate(s).
- Explore the FARS traffic accident data and create maps
- Create an R markdown file that contains all code to read the data, create the plots, and your answers.
- Teammates will review and revise each other's work.
- Submit!

## Submission Requirements

- A team member should create a GitHub repository called `ds202_lab5` which includes the group work.
- If you finish the lab individually, you only need to include your answers in `individual1.Rmd`
- If you collaborate, the team will work on `team.Rmd` to answer all the questions (for the final report). Also include your individual exploration in `individual1.Rmd` or `individual2.Rmd`, which will be your "scratch files". *Only the report generated from `team.Rmd` will be graded.*
- Include the data you used in your repository.
- The correct folder structure is as below. The folder `ds202_lab5` is a git repository.

    ```
    ds202_lab5
    |-- individual1.Rmd
    |-- (individual2.Rmd)
    |-- (team.Rmd)
    |-- *.html
    |-- Data files
    ```
    
## Background: FARS data

- US Department of transportation is keeping a record of every accident that results in a fatality in the FARS Data base (fatal accident report system, http://www.nhtsa.gov/FARS).
- FARS consists of 20+ tables consisting of various aspects of each accident
- Use the User's manual to find what the columns stand for. The manual can be found  at
https://crashstats.nhtsa.dot.gov/Api/Public/ViewPublication/813023
- Two of the main tables we are going to look at are the accident and the person tables. These data are made available here: [`accident`](https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv), and [`person`](https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv)
- Load the two datasets. 

    ```{r}
    acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
    # repeat for the person data
    ```

## Questions 

<!-- 1. Are there some days of the week where more accidents happen than the others (see FARS manual, use variable `DAY_WEEK`)? -->

1. Create a data frame containing the persons who are fatally hurt in the accidents (see FARS manual and look up variable `INJ_SEV`)
2. Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable `MAKE`)
3. Create a map, and label each state with the most dangerous vehicle. Discuss the definition of *the most dangerous* vehicle, and what you find from the map. (Hint: Read the description for the `STATE` and `COUNTY` columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from  [the General Services Administration’s (GSA) publication](https://www.gsa.gov/reference/geographic-locator-codes/glcs-for-the-us-and-us-territories). Use `readxl::read_xlsx` to read in the GLCs.)
4. Join the accident and person table (work out which variable(s) to use)
5. Tally the number of accidents by day of the week (DAY_WEEK), hour of the day (HOUR) and gender (SEX). Visualize the results and explain what you find.
6. Now plot a choropleth map of the number of deaths on a county level. Also explain what you find.
7. (Optional) Is summer or winter more dangerous? Does this depend on states? Explore and explain.


```{r,include=FALSE}
library(tidyverse)
acc=read.csv('fars2017/accident.csv')
dri=read.csv('fars2017/person.csv') %>% filter(PER_TYP == 1)
veh=read.csv('fars2017/vehicle.csv')
intersect(names(acc), names(dri))
a = acc %>% inner_join(dri)
```

